---
title: "Getting started with VisitationR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with VisitationR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5
)
```

```{r setup, include = FALSE}
library(VisitationR)
```

## The model

VisitationR is built around one assumption: seed set increases with pollinator
visits but saturates. Formally,

$$
\mathrm{seedset} = \frac{b\,a}{100} + \left(b - \frac{b\,a}{100}\right)\left(1 - e^{-c \cdot \mathrm{visits}}\right)
$$

with three parameters:

| Parameter | Meaning                                                        |
|-----------|----------------------------------------------------------------|
| `a`       | Autonomous selfing, as a percentage of the potential seed set   |
| `b`       | Potential (maximum) seed set                                    |
| `c`       | Per-visit saturation parameter, i.e. the pollination rate       |

The asymptote of the curve is `b`, and the value at zero visits is `b * a / 100`.
`c` is the quantity of ecological interest: it tells you how quickly a flower
moves from unpollinated towards fully pollinated.

## Fitting `c` to observational data

When you have counted visits and measured seed set in the field, use
`fit_data()`:

```{r}
visits <- c(0, 2, 5, 10, 20, 40, 80)
seedset <- c(20, 45, 70, 88, 96, 99, 100)

fit <- fit_data(seedset = seedset, visitation = visits)
fit
```

Ask for the standard errors instead of just the point estimates:

```{r}
fit_data(seedset = seedset, visitation = visits, simplify = "stats")
```

Keep the fitted object to compute goodness of fit and to plot:

```{r}
model <- fit_data(seedset = seedset, visitation = visits, simplify = "model")

pseudo_R2(model = model, seedset = seedset)
```

```{r}
plot_visits(
  a = coef(model)[["a"]],
  b = coef(model)[["b"]],
  c = coef(model)[["c"]],
  to_ = 90
)
```

## Deriving `c` from a single-visit experiment

If instead you measured three summary quantities — autonomous selfing (`a`),
potential seed set (`b`) and single-visit deposition (`SVD`) — you can obtain
`c` without running a field experiment. `calculate_visits()` discounts the
autonomous selfing baseline from `SVD`, simulates the decaying deposition
series, and fits the curve to it:

```{r}
my_c <- calculate_visits(a = 20, b = 100, SVD = 50)
my_c
```

The `optimal_loss` element is the per-visit decay rate, which you can also
obtain on its own with `loss()`.

## Inverting the model: how many visits do you need?

This is the usual question when designing an experiment or planning a
restoration. `calculate_required_visits()` answers "how many visits are needed
to reach 95% of the potential seed set?":

```{r}
calculate_required_visits(c_val = my_c$c_parameter, b = 100, a = 20)
```

Vary the target to see the whole trade-off:

```{r}
sapply(
  seq(0.5, 0.99, by = 0.05),
  function(p) calculate_required_visits(my_c$c_parameter, b = 100, a = 20,
                                       target_percent = p)
)
```

Note the asymmetry: going from 50% to 95% seed set requires many times more
visits than going from 0% to 50%. This is exactly why a small increase in
visitation can matter a lot in a pollination-limited system, and why the
assumption of saturation needs to be checked.

## Overlaying two populations

`plot_visits()` can be layered, which is handy for comparing a manipulated and a
control population:

```{r}
plot_visits(a = 30, b = 200, c = 10, to_ = 20, col_ = 2, lwd = 2)
plot_visits(a = 10, b = 150, c = 3, to_ = 20, col_ = 4, lty = 2, lwd = 2)
legend("bottomright", c("pollinated", "control"), col = c(2, 4),
       lty = c(1, 2), lwd = 2, bty = "n")
```

## From a transect to flowers visited

If visits were recorded along a fixed-area transect rather than per flower,
`V_transect_to_flower()` rescales the counts using the local flower density and
flower lifespan:

```{r}
V_transect_to_flower(V_transect = 50, flw_x_m2 = 100, lifespan = 8)
```

```{r sessioninfo}
sessionInfo()
```